---
title: "Summary Figure for Statistical Analyses in Australian Cohort and MyCode for 'Testing the performance of polygenic scores for multiple traits to explain cerebral palsy risk in two independent cohorts'"
author: "Jodi T Thomas"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    collapse: no
    df_print: paged
    highlight: textmate
    fig_caption: yes
    fig_height: 6
    fig_width: 12
    toc: yes
    toc_float: yes
    toc_depth: 4
    code_folding: hide
  pdf_document:
    df_print: default
    fig_caption: yes
    fig_height: 6
    fig_width: 12
    highlight: tango
    latex_engine: xelatex
    number_sections: yes
    toc: yes
    toc_depth: 2
  word_document:
    fig_caption: yes
    fig_height: 6
    fig_width: 12
    highlight: tango
    toc: yes
    toc_depth: 2
documentclass: article
classoption: a4paper
fontsize: 12pt
mainfont: Arial
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval=FALSE, warning=FALSE, message=FALSE, error=FALSE)
options(tinytex.engine = 'xelatex')
```

```{r Setwd, results='markdown', eval=TRUE, echo=FALSE}
#Set working directory to location of folder holding all files
setwd("/path/R_Aus_CP_Biobank_PGS_regressions/03_Summary_figure/")
```

```{r readRScripts, results='markdown', eval=TRUE, echo=FALSE}
#Read in R scripts
knitr::read_chunk('Scripts/01_Summary_figure.R')
```

```{r LoadDataRmd, eval = TRUE, echo = FALSE}
# Full dataset results: Aus + MyCode
load(file = "Results/Aus_MyCode_PGS_results_table.RData")

``` 

# Load Packages
```{r LoadPackages, eval=TRUE}
```

# Summary Figure

## Dataframe of all results

```{r FullDataframe}

```

## Figure: Australian cohort and MyCode 

* Results from full Australian cohort and Matched MyCode cohort
* Using adjusted models for logistic regression

```{r SummaryFigureFull, eval = TRUE, fig.height=15, fig.width=11}

```

Save Figure

```{r SaveFigureFull}

```

## Figure: Australian cohort stratified by monogenic diagnosis

* Results from full Australian cohort all data or stratified by monogenic diagnosis
* Using adjusted models for logistic regression

```{r SummaryFigureAusStrat, eval = TRUE, fig.height=15, fig.width=11}

```

Save Figure

```{r SaveFigureAusStrat}

```

## Figure: All datasets checking age effect

* Results from Australian cohort full data or controls restricted to youngest 525 participants and MyCode all data or MyCode data matched on age
* Using adjusted models for logistic regression

```{r SummaryFigureFullAndCheckAge, eval = TRUE, fig.height=15, fig.width=11}

```

Save Figure

```{r SaveFigureFullAndCheckAge}

```


# SessionInfo
```{r eval = TRUE}
sessionInfo()
```