---
title: "Statistical Analyses in Australian Cohort for 'Testing the performance of polygenic scores for multiple traits to explain cerebral palsy risk in two independent cohorts'"
author: "Jodi T Thomas"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    collapse: no
    df_print: paged
    highlight: textmate
    fig_caption: yes
    fig_height: 6
    fig_width: 12
    toc: yes
    toc_float: yes
    toc_depth: 4
    code_folding: hide
  pdf_document:
    df_print: default
    fig_caption: yes
    fig_height: 6
    fig_width: 12
    highlight: tango
    latex_engine: xelatex
    number_sections: yes
    toc: yes
    toc_depth: 2
  word_document:
    fig_caption: yes
    fig_height: 6
    fig_width: 12
    highlight: tango
    toc: yes
    toc_depth: 2
documentclass: article
classoption: a4paper
fontsize: 12pt
mainfont: Arial
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval=FALSE, warning=FALSE, message=FALSE, error=FALSE)
options(tinytex.engine = 'xelatex')
```

```{r Setwd, results='markdown', eval=TRUE, echo=FALSE}
#Set working directory to location of folder holding all files
setwd("/path/Aus_CP_Biobank/R_Aus_CP_Biobank_PGS_regressions/01_Australian_cohort/")
```

```{r readRScripts, results='markdown', eval=TRUE, echo=FALSE}
#Read in R scripts
knitr::read_chunk('Scripts/01_CP_Data.R')
knitr::read_chunk('Scripts/02_Sample_sizes.R')
knitr::read_chunk('Scripts/03_Check_PGS.R')
knitr::read_chunk('Scripts/04_Exploration.R')
knitr::read_chunk('Scripts/05_Fit_models.R')
knitr::read_chunk('Scripts/06_Check_models.R')
knitr::read_chunk('Scripts/07_Regression_results.R')
knitr::read_chunk('Scripts/08_AUC_results.R')
knitr::read_chunk('Scripts/09_R2_liability_results.R')
```

```{r LoadDataRmd, eval = TRUE, echo = FALSE}

# Relatedness data
related <- read.table("/path/Cerebral_palsy/Aus_CP_Biobank/Relatedness/Aus_CPBiobank_QSKIN_relatedness.genome", header = T)

# Dataframe
load(file = "Data/CP_df_not_related.Rdata")
load(file = "Data/Monogenic_only_df.RData")
load(file = "Data/Non_monogenic_df.RData")
load(file = "Data/Young525_df.RData")

# Lists of PGS variables and names
load(file = "Data/PGS_variables.RData")
load(file = "Data/PGS_names.RData")

# Exploration
load(file = "Results/Exploration.RData")

# Regression Models
load(file = "Results/Regression_models_with_covar.RData")
load(file = "Results/Regression_models_no_covar.RData")
# Regression dataframe
load(file = "Results/regression_with_covar_df.RData")
load(file = "Results/regression_no_covar_df.RData")

# ROC Models
load(file = "Results/roc_models.RData")
# ROC dataframe
load(file = "Results/roc_df.RData")

# R2 liability
load(file = "Results/r2l_df.RData")

``` 


# Load Packages
```{r LoadPackages, eval=TRUE}
```


# Set Up Data

## Read in the data
```{r LoadData}

```

## Tidy Data

* Genetic_Sex = sex of individual determined by genetic data. 1 = male, 2 = female
* CP_all = presence of cerebral palsy. 1 = case, 0 = control
  + Cases = all people in Aus CP biobank
  + Controls = all people from QSKIN I and II
* PC1 - PC20 = ancestry PCs
* age_sample = age at when sample was given (saliva for QSKIN, blood for Aus CP Biobank)
* FID = Family ID (CP_Adelaide_GSA for all individuals from Aus CP Biobank)
* IID = Individual ID used for genetic data
* CP_ID = ID used in Aus CP Biobank (not the genetic ID). FALSE for QSKIN individuals
* PGS_trait_CT = PGS of trait using clumping and thresholding with all SNPs
* PGS_trait_SBayesRC = PGS of trait using SBayesRC

* Only in CP Aus Biobank individuals
  + gestation = gestation duration in weeks
  + birthwtgrams = birth weight in grams
  + gmfcs = The gross motor classification function score. 1-5 or unknown
  + genetic = confirmed monogenic diagnosis for CP in that individual by whole exome or whole genome sequencing. yes or no or unknown
  + MRIorUSabnormal = Abnormal brain damage pattern found on MRI, CT or Ultrasound: yes or no. 'Unknown' is where we have no data on whether neuroimaging was done, whereas  'no imaging' indicates no neuroimaging has been done for that patient.
    
* Only in QSKIN individuals
  + Have you ever been diagnosed with, experienced or been treated for any of the following conditions? 0 = no, 1 = yes
    + EPILEPSY_QSKIN
    + STROKE_QSKIN
    + ADHD_QSKIN
    + AUTISM_QSKIN

```{r TidyData}

```


## Relatedness
```{r Relatedness, eval = TRUE}

```

## Remove relatedness

* Remove 1 individual from each pair that have pi_hat > 0.09375 with controls preferentially removed.
  + This removes 3rd degree relatives (threshold is halfway between 3rd and 4th degree relatives)

```{r RemoveRelatedness}

```



## Stratify Monogenic Diagnosis

* Stratify dataset based on presence or absence of monogenic diagnosis for CP

```{r StratDF}

```

## Restrict controls to youngest participants

```{r AgeDF}

```


# Sample Sizes {.tabset .tabset-fade}

## Full dataset
```{r SampleSizesFullDF, eval = TRUE}

```

## Cases with a monogenic diagnosis
```{r SampleSizesMonoDF, eval = TRUE}

```

## Cases without a monogenic diagnosis
```{r SampleSizesNonMonoDF, eval = TRUE}

```

## Controls restricted to youngest 525 participants
```{r SampleSizesYoung525DF, eval = TRUE}

```

## Age Distributions
```{r CheckAgeDistribution, eval = TRUE}

```


# Check PGS

* Check each PGS associates with appropriate self-report yes/no in QSKIN

## Epilepsy

```{r CheckPGSEpilepsy, eval = TRUE}

```

## Stroke

```{r CheckPGSStroke, eval = TRUE}

```

## Autism

```{r CheckPGSAutism, eval = TRUE}

```

## Cerebral Palsy
```{r CheckPGSCP, eval = TRUE}

```

## Birth Weight
```{r CheckPGSBirthWeight, eval = TRUE}

```



# Data Exploration {.tabset .tabset-fade}

Set up data exploration
```{r SetUpPGSVars}

```

```{r Exploration}

```



## Full dataset

### Distribution of PGS (raw and standardised): Histograms

```{r FullExplorePGSDistributionRawStand, eval = TRUE}

```

### Distribution of PGS (raw and standardised): Summary Table

```{r FullExplorePGSDistributionRawStandSummaryTable, eval = TRUE}

```

### Distribution of PGS grouped by CP status

```{r FullExploreDistributionbyCP, eval = TRUE, fig.height=12}

```

## Monogenic diagnosis only

### Distribution of PGS (raw and standardised): Histograms

```{r MonoExplorePGSDistributionRawStand, eval = TRUE}

```

### Distribution of PGS (raw and standardised): Summary Table

```{r MonoExplorePGSDistributionRawStandSummaryTable, eval = TRUE}

```

### Distribution of PGS grouped by CP status

```{r MonoExploreDistributionbyCP, eval = TRUE, fig.height=12}

```

## No monogenic diagnosis

### Distribution of PGS (raw and standardised): Histograms

```{r NonmonoExplorePGSDistributionRawStand, eval = TRUE}

```

### Distribution of PGS (raw and standardised): Summary Table

```{r NonmonoExplorePGSDistributionRawStandSummaryTable, eval = TRUE}

```

### Distribution of PGS grouped by CP status

```{r NonmonoExploreDistributionbyCP, eval = TRUE, fig.height=12}

```


## Youngest 525 controls

### Distribution of PGS (raw and standardised): Histograms

```{r Young525ExplorePGSDistributionRawStand, eval = TRUE}

```

### Distribution of PGS (raw and standardised): Summary Table

```{r Young525ExplorePGSDistributionRawStandSummaryTable, eval = TRUE}

```

### Distribution of PGS grouped by CP status

```{r Young525ExploreDistributionbyCP, eval = TRUE, fig.height=12}

```

# Fit Models

## Logistic Regression {.tabset .tabset-fade}

### Adjusted Models

```{r RegrModelsWithCovar}

```

### Unadjusted Models
```{r RegrModelsNoCovar}

```


## Receiver Operating Characteristic Curve


```{r ROCFitPGS}

```


# Check Models {.tabset .tabset-fade}

* PC5 and PC6 have high VIF, but as these are only as covariates to be adjusted for and not interested in their estimates OK

## Full dataset: Adjusted Models

### One PGS in each model

#### Check_model
```{r FullModelCheckPGSModelsWithCovarCheckModel}

```

#### DHARMa
```{r FullModelCheckPGSModelsWithCovarDHARMa}

```

### All PGS in same model
```{r FullModelWithCovarCheckAllPGSModel}

```

## Full dataset: Unadjusted Models

### One PGS in each model

#### Check_model
```{r FullModelCheckPGSModelsNoCovarCheckModel}

```

#### DHARMa
```{r FullModelCheckPGSModelsNoCovarDHARMa}

```

### All PGS in same model
```{r FullModelNoCovarCheckAllPGSModel}

```

## Monogenic diagnosis only: Adjusted Models

### One PGS in each model

#### Check_model
```{r MonoModelWithCovarCheckPGSModelsCheckModel}

```

#### DHARMa
```{r MonoModelWithCovarCheckPGSModelsDHARMa}

```

### All PGS in same model
```{r MonoModelWithCovarCheckAllPGSModel}

```

## Monogenic diagnosis only: Unadjusted Models

### One PGS in each model

#### Check_model
```{r MonoModelNoCovarCheckPGSModelsCheckModel}

```

#### DHARMa
```{r MonoModelNoCovarCheckPGSModelsDHARMa}

```

### All PGS in same model
```{r MonoModelNoCovarCheckAllPGSModel}

```


## No monogenic diagnosis: Adjusted Models

### One PGS in each model

#### Check_model
```{r NonmonoModelWithCovarCheckPGSModelsCheckModel}

```

#### DHARMa
```{r NonmonoModelWithCovarheckPGSModelsDHARMa}

```

### All PGS in same model
```{r NonmonoModelWithCovarCheckAllPGSModel}

```


## No monogenic diagnosis: Unadjusted Models

### One PGS in each model

#### Check_model
```{r NonmonoModelNoCovarCheckPGSModelsCheckModel}

```

#### DHARMa
```{r NonmonoModelNoCovarheckPGSModelsDHARMa}

```

### All PGS in same model
```{r NonmonoModelNoCovarCheckAllPGSModel}

```


## Youngest 525 Controls: Adjusted Models

### One PGS in each model

#### Check_model
```{r Young525ModelWithCovarCheckPGSModelsCheckModel}

```

#### DHARMa
```{r Young525ModelWithCovarheckPGSModelsDHARMa}

```

### All PGS in same model
```{r Young525ModelWithCovarCheckAllPGSModel}

```

## Youngest 525 Controls: Unadjusted Models

### One PGS in each model

#### Check_model
```{r Young525ModelNoCovarCheckPGSModelsCheckModel}

```

#### DHARMa
```{r Young525ModelNoCovarheckPGSModelsDHARMa}

```

### All PGS in same model
```{r Young525ModelNoCovarCheckAllPGSModel}

```




# Regression Results {.tabset .tabset-fade}

```{r ResultsRegr}

```

## Full dataset: Adjusted Models
```{r FullResultsWithCovarRegrPrint, eval = TRUE}

```

## Full dataset: Unadjusted Models
```{r FullResultsNoCovarRegrPrint, eval = TRUE}

```

## Monogenic diagnosis only: Adjusted Models
```{r MonoResultsWithCovarRegrPrint, eval = TRUE}

```

## Monogenic diagnosis only: Unadjusted Models
```{r MonoResultsNoCovarRegrPrint, eval = TRUE}

```

## No monogenic diagnosis: Adjusted Models
```{r NonmonoResultsRegrPrint, eval = TRUE}

```

## No monogenic diagnosis: Unadjusted Models
```{r NonmonoResultsNoCovarRegrPrint, eval = TRUE}

```

## Youngest 525 controls: Adjusted Models
```{r Young525ResultsWithCovarRegrPrint, eval = TRUE}

```

## Youngest 525 controls: Unadjusted Models
```{r Young525ResultsNoCovarRegrPrint, eval = TRUE}

```


# Receiver Operating Characteristic Curve Results

* Area Under Receiver Operator Characteristic Curve (AUC)
    + AUC = Probability that a randomly selected case has a higher test score than a randomly selected control
    + Well established measure of validity of tests for classifier diseased vs nondiseased individuals
    + Nice property  independent to proportion of cases and controls in sample
    + Range 0.5 to 1
    + 0.5 = the the PGS does not discriminate cases from controls, 1 = perfect discrimination


## Results {.tabset .tabset-fade}

```{r AUCResults}

```

### Full dataset

#### Results
```{r FullResultsAUCPrint, eval = TRUE}

```

#### AUC Plots

```{r FullAUCPlot, eval = TRUE}

```

### Monogenic diagnosis only

#### Results
```{r MonoResultsAUCPrint, eval = TRUE}

```

#### AUC Plots
```{r MonoAUCPlot, eval = TRUE}

```

### No monogenic diagnosis

#### Results
```{r NonmonoResultsAUCPrint, eval = TRUE}

```

#### AUC Plots
```{r NonmonoAUCPlot, eval = TRUE}

```

### Youngest 525 controls

#### Results
```{r Young525ResultsAUCPrint, eval = TRUE}

```

#### AUC Plots
```{r Young525AUCPlot, eval = TRUE}

```

# Variance Explained on the Liability Scale

* Variance in CP risk explained on the liability scale, attributable to each PGS
* Linear models
  + Full model: CP(case/control) ~ covariates + score
  + Reduced model: CP(case/control) ~ covariates
* Convert observed R2 from each model to the liability scale using CP prevalence of 0.0014 (1 in 700 in Australia)
  + Use R code from Lee SH, et al., A Better Coefficient of Determination for Genetic Profile Analysis. Genetic Epidemiology, 2012. 36(3):214-224.
* R2 liability of full model - R2 liability of reduced model = R2 liability attributable to PGS

## Function to convert observed to liability scale
```{r VarExplainedLiabilityScaleFunction}

```

## Caluclate R^2 (liability)
```{r VarExplainedLiabilityScale}

```

## Results {.tabset .tabset-fade}

### Full dataset

```{r FullR2lPrint, eval = TRUE}

```

### Monogenic diagnosis only

```{r MonoR2lPrint, eval = TRUE}

```

### No monogenic diagnosis

```{r NonmonoR2lPrint, eval = TRUE}

```

### Youngest 525 controls

```{r Young525R2lPrint, eval = TRUE}

```



# SessionInfo
```{r eval = TRUE}
sessionInfo()
```

